rbcL ASV clustering

### 2019-11-05 code/rbcl_asv_clustering.Rmd
aesthetics

```{r echo=F, fig.width=8, fig.height=6}
ggplot(tsne_added_names, aes(x=V1, y=V2)) +
  ggforce::geom_mark_ellipse(aes(fill=otu_name, label=otu_name), alpha=0.3, show.legend = FALSE) +
  # plot points not assigned to genus or species
  geom_point(data = filter(tsne_added_names, is.na(plot_taxa)), 
             color = "grey",alpha = 0.8, size = 3) +
  # color points that were assigned to genus or species
  geom_point(data = filter(tsne_added_names, !is.na(plot_taxa)), 
             aes(color = plot_taxa %>% fct_relevel(taxa.order)), alpha = 0.8, size = 3) +
  labs(title = "Points colored by both genus and species", color="ASV taxonomic assignment") +
  theme_few() 
```

```{r include_rel_abund, echo=F}
total_rel_abund <- rel_abund %>% 
  group_by(asv_id) %>% 
  summarise(global_rel_abund = sum(rel_abund))

tsne_rel_abund <- tsne_added_names %>% 
  left_join(total_rel_abund, by="asv_id")

final_tsne <- ggplot(tsne_rel_abund, aes(x=V1, y=V2)) +
  ggforce::geom_mark_ellipse(aes(fill=otu_name, label=otu_name), 
                             show.legend = FALSE, 
                             alpha=0.2, 
                             linetype="blank", 
                             expand = unit(2, "mm")) +
  # plot points not assigned to genus or species
  geom_point(data = filter(tsne_rel_abund, is.na(plot_taxa)), 
             aes(size = global_rel_abund),
             color = "grey",alpha = 0.5) +
  # color points that were assigned to genus or species
  geom_point(data = filter(tsne_rel_abund, !is.na(plot_taxa)), 
             aes(color = plot_taxa %>% fct_relevel(taxa.order),
                 size = global_rel_abund), 
             alpha = 0.5) +
  labs(color="ASV taxonomic assignment", tag="A") +
  ggthemes::theme_few() +
  scale_size(range = c(1, 6), guide="none") + ################################################# not changing the scale size
  theme(legend.position = "none")
final_tsne
```

```{r asv_mds, echo=F}
mds_coords <- dists %>% 
  cmdscale() %>% # ok to use euclidian dist, no null values when comparing relatedness
  as_tibble() %>% 
  add_column(asv_id = colnames(dists), .before=1)

@@

mds <- ggplot(mds_asv_plot, aes(V1, V2)) + 
  ggforce::geom_mark_ellipse(data = mds_asv_plot %>% 
                               filter(otu_name != "Trebouxiophyceae J"), # dont plot the ellipse around other treb
                             aes(fill = otu_name, label=otu_name),
                             show.legend = FALSE, 
                             alpha=0.2, 
                             linetype="blank",
                             expand = unit(2, "mm")) +
  geom_point(aes(alpha = 0.7, color = plot_taxa, size = global_rel_abund)) +
  scale_size(range = c(1, 6), guide="none") +
  scale_alpha(guide="none")+
  theme_few() +
  labs(color="ASV taxonomic assignment", tag="B")
  # geom_text(aes(label = asv_id)) # too see outliers, asv_id=c("639", "534")
mds
```

@@

ggsave(filename = here("manuscript/figs/2_taxonomy/tsne.png"), plot=final_tsne, height = 6, width = 8)
ggsave(filename = here("manuscript/figs/2_taxonomy/mds.png"), plot=mds,  height = 6, width = 8)

>>>>

```{r echo=F, fig.width=10, fig.height=10}
ggplot(tsne_added_names, aes(x=V1, y=V2)) +
  ggforce::geom_mark_ellipse(aes(fill=otu_name, label=otu_name), alpha=0.3, show.legend = FALSE) +
  # plot points not assigned to genus or species
  geom_point(data = filter(tsne_added_names, is.na(plot_taxa)), 
             color = "grey",alpha = 0.8, size = 3) +
  # color points that were assigned to genus or species
  geom_point(data = filter(tsne_added_names, !is.na(plot_taxa)), 
             aes(color = plot_taxa %>% fct_relevel(taxa.order)), alpha = 0.8, size = 3) +
  labs(title = "Points colored by both genus and species", color="ASV taxonomic assignment") +
  theme_few() 
```

```{r include_rel_abund, echo=F, fig.width=5, fig.height=4}
total_rel_abund <- rel_abund %>% 
  group_by(asv_id) %>% 
  summarise(global_rel_abund = sum(rel_abund))

tsne_rel_abund <- tsne_added_names %>% 
  left_join(total_rel_abund, by="asv_id") %>% 
  # remove C kri ellipse
  mutate(otu_name = if_else(otu_name == "Chloromonas krienitzii", "Other Chloromonas", otu_name))

final_tsne <- ggplot(tsne_rel_abund, aes(x=V1, y=V2)) +
  ggforce::geom_mark_ellipse(aes(fill=otu_name, label=otu_name), 
                             show.legend = FALSE, 
                             alpha=0, 
                             linetype="dashed", 
                             label.margin = margin(1,1,1,1,"mm"),
                             label.buffer = unit(1, "mm"),
                             label.fontface = "plain",
                             label.fontsize = 8,
                             con.type = 0,
                             expand = unit(2, "mm")) +
  geom_point(aes(color = plot_taxa %>% fct_relevel(taxa.order), size = global_rel_abund, alpha=0.5))+
  labs(color="ASV taxonomic assignment", tag="A", title="t-SNE") +
  ggthemes::theme_few() +
  scale_size(range = c(2, 12), guide="none")

tsne_no_legend <- final_tsne +
    theme(legend.position = "none")
tsne_no_legend
```

```{r, fig.height=4, fig.width=2.2}
p <- final_tsne +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=3)), alpha="none") 
legend <- p %>% 
  get_legend() %>% 
  as_ggplot()
legend
```

```{r asv_mds, echo=F, fig.width=5, fig.height=4}
mds_coords <- dists %>% 
  cmdscale() %>% # ok to use euclidian dist, no null values when comparing relatedness
  as_tibble() %>% 
  add_column(asv_id = colnames(dists), .before=1)

@@

mds <- ggplot(mds_asv_plot, aes(V1, V2)) + 
  ggforce::geom_mark_ellipse(data = mds_asv_plot %>% 
                               filter(otu_name != "Trebouxiophyceae J"), # dont plot the ellipse around other treb
                             aes(fill = otu_name, label=otu_name),
                             show.legend = FALSE, 
                             alpha=0, 
                             linetype="dashed", 
                             label.margin = margin(1,1,1,1,"mm"),
                             label.buffer = unit(1, "mm"),
                             label.fontface = "plain",
                             label.fontsize = 8,
                             con.type = 0,
                             expand = unit(2, "mm")) +
  geom_point(aes(color = plot_taxa %>% fct_relevel(taxa.order), size = global_rel_abund, alpha=0.5))+
  labs(color="ASV taxonomic assignment", tag="B",title="MDS") +
  ggthemes::theme_few() +
  scale_size(range = c(2, 12), guide="none")+
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=3)), alpha="none") +
  theme(legend.position = "none")

  # geom_point(aes(alpha = 0.7, color = plot_taxa, size = global_rel_abund)) +
  # scale_size(range = c(1, 6), guide="none") +
  # scale_alpha(guide="none")+
  # theme_few() +
  # labs(color="ASV taxonomic assignment", tag="B")
  # geom_text(aes(label = asv_id)) # too see outliers, asv_id=c("639", "534")
mds
```

```{r, fig.width=4, fig.height=6}
asv_arrange <- grid.arrange(tsne_no_legend, mds, nrow=2)
```

@@

ggsave(filename = here("manuscript/figs/2_taxonomy/asvs.pdf"), plot=asv_arrange, height = 6, width = 4)
ggsave(filename = here("manuscript/figs/2_taxonomy/asv_legend.pdf"), plot=legend,  height = 4, width = 2.2)
